---
title: "R Project"
author: 'Yoni Kornblau'
---

```{r}
library(dplyr)
library(ggplot2)
library(shiny)
library(tidyverse)
data = read.csv('Medicare_Inpatient_Hospital_by_Provider_and_Service_2019.csv')
head(data)
```
Cleaning the data: 
```{r}
better_data =data %>% 
  rename(
    CCN = ï..Rndrng_Prvdr_CCN,
    Name = Rndrng_Prvdr_Org_Name,
    City = Rndrng_Prvdr_City,
    State = Rndrng_Prvdr_State_Abrvtn,
    Address = Rndrng_Prvdr_St,
    FIPS = Rndrng_Prvdr_State_FIPS,
    Zip_code = Rndrng_Prvdr_Zip5,
    RUCA = Rndrng_Prvdr_RUCA,
    RUCA_desc = Rndrng_Prvdr_RUCA_Desc,
    Diagnosis_code = DRG_Cd,
    Diagnosis_desc = DRG_Desc
    )
```

```{r}
summary(better_data)
better_data$Avg_Submtd_Cvrd_Chrg=str_sub(better_data$Avg_Submtd_Cvrd_Chrg,2)
better_data$Avg_Submtd_Cvrd_Chrg = as.numeric(sub(',','',better_data$Avg_Submtd_Cvrd_Chrg))
better_data$Avg_Tot_Pymt_Amt=str_sub(better_data$Avg_Tot_Pymt_Amt,2)
better_data$Avg_Tot_Pymt_Amt = as.numeric(sub(',','',better_data$Avg_Tot_Pymt_Amt))
better_data$Avg_Mdcr_Pymt_Amt=str_sub(better_data$Avg_Mdcr_Pymt_Amt,2)
better_data$Avg_Mdcr_Pymt_Amt = as.numeric(sub(',','',better_data$Avg_Mdcr_Pymt_Amt))
better_data$Tot_Dschrgs = as.numeric(better_data$Tot_Dschrgs)
better_data = better_data %>% mutate(Pct_Chrgd_Paid = Avg_Tot_Pymt_Amt/Avg_Submtd_Cvrd_Chrg)
better_data = better_data %>% mutate(OOP_Amt_Paid = Avg_Tot_Pymt_Amt - Avg_Mdcr_Pymt_Amt)
head(better_data)

```



```{r}
n_distinct(better_data$Name)
n_distinct(better_data$CCN)
n_distinct(better_data$Diagnosis_code)
sum_discharges = better_data$Tot_Dschrgs %>% replace_na(0) %>% as.numeric() %>% sum()
unique(better_data$Diagnosis_desc) %>% sort()
n_distinct(better_data$City)
unique(better_data$RUCA_desc)

```
```{r}

Tot_pmnt_graph = better_data %>% arrange(desc(Avg_Tot_Pymt_Amt)) %>% pull(Avg_Tot_Pymt_Amt) %>% plot()
Submtd_cvrd_chrg_graph = better_data %>% arrange(desc(Avg_Submtd_Cvrd_Chrg)) %>% pull(Avg_Submtd_Cvrd_Chrg) %>% plot()
Mdcr_pmnt_graph = better_data %>% arrange(desc(Avg_Mdcr_Pymt_Amt)) %>% pull(Avg_Mdcr_Pymt_Amt) %>% plot()
```
```{r}
better_data %>% arrange(desc(Avg_Mdcr_Pymt_Amt)) %>% head(1000) %>% ggplot(aes(x=Avg_Mdcr_Pymt_Amt,))+geom_bar()
```
Create column of amt_patient_paid = total_payment - medicare_payment
Create column percent of charge_paid= charged amt/ total_payment
if time: create column where all the cc/mcc is removed
```{r}
better_data %>% mutate(Pct_chrgd_paid = Avg_Tot_Pymt_Amt/Avg_Submtd_Cvrd_Chrg)
better_data %>% mutate(OOP_amt_paid = Avg_Tot_Pymt_Amt - Avg_Mdcr_Pymt_Amt)
better_data %>% mutate(Diagnosis_desc = )
```


Graphs to make:
1. Top most costly procedures by charged_amt / least costly
2. Does amt_patient_paid/total_payment change based off type of procedure / hospital
3. Graph location of hospitals on US map with coloring of RUCA codes
4. Hospital by number of patients
5. procedure by total number of patients
6. Graph with all payment types overlaid
7. glm for predicting price of procedure using zip code, hospital, procedure type
8. Does percent of charge paid vary by procedure/hospital?



```{r}
plot(world)
```
```{r}
library(zipcodeR)
zips = geocode_zip(better_data$Zip_code) %>% drop_na()
us = map_data('state')
ggplot(zips, aes(y = lat,x = lng)) + geom_polygon(data= us, aes(x=long,y=lat), color='blue',fill=NA,alpha=.35)+
  geom_point(size=.15,alpha=.25) +
  xlim(-125,-65)+ylim(20,50)

```

